CUNY SPS DATA 605 Final

### 2019-05-24 DATA605/MSilva_Final.Rmd
More work on the final

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(Matrix)
library(matlib)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

*small letter "x" is `r x` and small letter "y" is `r y`.*

@@

*There are some numerical features that are really categories that will need to be cleaned up*

```{r}
clean_MSSubClass <- function(x){
  return(paste0("SC", as.character(x)))
}

clean_MoSold <- function(x){
  if(x > 9){return(as.character(x))}
  else{
    return(paste0("0", as.character(x)))
  }
}

clean_text_number <- function(x){
  if(x < 10){return(paste0("0", x))}
  return(as.character(x))
}

kaggle <- kaggle %>%
  # Clean up variables
  rowwise() %>%
  mutate(MSSubClass = clean_MSSubClass(MSSubClass)) %>%
  mutate(MoSold = clean_MoSold(MoSold)) %>%
  mutate(OverallQual = clean_text_number(OverallQual)) %>%
  mutate(OverallCond = clean_text_number(OverallCond)) %>%
  ungroup() %>%
  mutate(MSSubClass = as.factor(MSSubClass),
         MoSold = as.factor(MoSold),
         OverallQual = as.factor(OverallQual),
         OverallCond = as.factor(OverallCond)) %>%
  as.data.frame()
```

@@

```{r echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
for (variable in  names(kaggle)){
  cat(paste("<h4>", variable, "</h4>"))

@@

```{r, results="asis", tidy=TRUE}
zero_vars <- c()

@@

cat("<table><thead><tr><th>Variable</th><th>Correlation</th><th>Equal to Zero</th></tr></thread><tbody>")

for(variable in variables){
  d <- kaggle[,names(kaggle) == variable]
  if(is.numeric(d)){
    results <- cor.test(kaggle$SalePrice, d, conf.level = 0.8)
    if(0 > results$conf.int[1] & results$conf.int[2] > 0){
      hypothesis_test_results <- "Yes"
      zero_vars <- c(zero_vars, variable)
    } else {
      hypothesis_test_results <- "No"
    }
    cat(paste("<tr><td>",variable,"</td><td>", cor(kaggle$SalePrice, d), "</td><td>", hypothesis_test_results, "</td>"))
  }
}
cat("</tbody></table>")
```

@@

*The LU decomposition should yield the correlation matrix after multiplying the two components.*

@@

A %>%
  kable() %>%
  kable_styling()

@@

*The optimal value for $\lambda$ is `r lambda`.*
```{r}
samples <- rexp(10000, rate = lambda)
```

@@

```{r}
histogram_data <- data.frame(Value = samples, Data = c("Sample"))
histogram_data <- data.frame(Value = kaggle$BsmtUnfSF, Data = c("Original")) %>%
  rbind(histogram_data) %>%
  mutate(Data = as.factor(Data))

@@

*The 95% confidence interval (assuming normality) is `r round(ci[1])` to `r round(ci[2])`.*

@@

*The actual 95% confidence interval is `r quantile(kaggle$BsmtUnfSF, c(0.05))` to `r quantile(kaggle$BsmtUnfSF, c(0.95))`.*

@@

*Since there are so many variables I will be using a random forest to detect which variables are the most important and will be using them to build the model.*

```{r, cache=TRUE, message=FALSE, warning=FALSE}
library(randomForest)
library(caret)

set.seed(12345)

rf_data <- kaggle %>%
  # Dropping variables with a lot of missing values
  select(-Alley, -MasVnrType, -MasVnrArea, -BsmtQual, -BsmtCond, -BsmtExposure, -BsmtFinType1, -BsmtFinType2, -Electrical, -FireplaceQu, -GarageType, -GarageYrBlt, -GarageFinish, -GarageQual, -GarageCond, -PoolQC, -Fence, -MiscFeature)

ctrl <- trainControl(method = "cv", number = 10)
rf_fit <- train(SalePrice ~ ., data = rf_data, method = "rf", trControl = ctrl, importance = TRUE, verbose = TRUE)
```

*Now we can examine the variable importance*

```{r}
imp_var <- importance(rf_fit$finalModel) 
variables <- rownames(imp_var)

@@

>>>>

```{r, message=FALSE, warning=FALSE}
library(randomForest)
library(caret)
library(MASS)
library(Matrix)
library(matlib)
library(dplyr)
library(ggplot2)
library(tidyr)
library(kableExtra)
```

@@

*Small letter "x" is `r x` and small letter "y" is `r y`.*

@@

@@

```{r stats_and_plots, echo=FALSE, eval=TRUE, results='asis', warning=FALSE}
for (variable in  names(kaggle)){
  cat(paste("<h4>", variable, "</h4>"))

@@

```{r, results="asis"}
zero_vars <- c()

@@

cat("<table><thead><tr><th>Variable</th><th>Correlation Equal to Zero</th></tr></thread><tbody>")

for(variable in variables){
  d <- kaggle[,names(kaggle) == variable]
  if(is.numeric(d)){
    results <- cor.test(kaggle$SalePrice, d, conf.level = 0.8)
    if(0 > results$conf.int[1] & results$conf.int[2] > 0){
      hypothesis_test_results <- "Yes"
      zero_vars <- c(zero_vars, variable)
    } else {
      hypothesis_test_results <- "No"
    }
    cat(paste("<tr><td>",variable,"</td><td>", hypothesis_test_results, "</td>"))
  }
}
cat("</tbody></table>")
```

@@

*The LU decomposition should yield the correlation matrix after multiplying the two components.  In other words this:*

@@

A %>%
  kable() %>%
  kable_styling()
```

*Should match this:*

```{r}
correlation_matrix %>%
  kable() %>%
  kable_styling()
```

*Which it does.*

@@

*The optimal value for $\lambda$ is `r lambda`.*

```{r}
samples <- rexp(10000, rate = lambda)
```

@@

```{r, echo=FALSE}
histogram_data <- data.frame(Value = samples, Data = c("Sample"))
histogram_data <- data.frame(Value = kaggle$BsmtUnfSF, Data = c("Original")) %>%
  rbind(histogram_data) %>%
  mutate(Data = as.factor(Data))

@@

*The 95% confidence interval (assuming normality) is `r round(ci[1])` to `r round(ci[2])`. Although what I think was meant was assume the data is normally distributed.  Calculate the 5th and 95th percentile.  That is found using $x = \mu + Z \sigma$.  The Z is -1.645 for the 5th percentile, and 1.645 for the 95th.  So the percentiles would be `r round(mu + (-1.645 * s), 0)` and `r round(mu + (1.645 * s), 0)`.*

@@

*The actual 5th percentile is `r quantile(kaggle$BsmtUnfSF, c(0.05))` and the 95th is `r quantile(kaggle$BsmtUnfSF, c(0.95))`.  So the findings are summarized in the following table:*

| Method | 5% | 95% |
| ------ |:--:|:---:|
| Exponential CDF | `r round(-log(0.95)/lambda,0)` | `r round(-log(0.05)/lambda,0)` |
| Normal 95% CI | `r round(ci[1])`      | `r round(ci[2])` |
| Normal Percentiles | `r round(mu + (-1.645 * s), 0)`  | `r round(mu + (1.645 * s), 0)` | 
| Emperical Percentiles | `r round(quantile(kaggle$BsmtUnfSF, c(0.05)),0)`       | `r round(quantile(kaggle$BsmtUnfSF, c(0.95)),0)` |

*If we model the data as exponentially distributed the 5th percentile is `r round(-log(0.95)/lambda,0)`.  If we model it as normally distributed the 5th is at `r round(mu + (-1.645 * s), 0)` which in the context of square footage does not make any sense. The actual 5th percentile is `r round(quantile(kaggle$BsmtUnfSF, c(0.05)),0)`. The difference is explained to the assumed shape/distibution of the underlying data.*

*Looking at the 95th percentile we have `r round(-log(0.5)/lambda,0)` if the data are exponentially distributed, `r round(mu + (1.645 * s), 0)` if it is normally distributed and `r round(quantile(kaggle$BsmtUnfSF, c(0.95)),0)` in reality.  Again the difference is due to the assumed shape.*

*I have left out the 95% CI from the discussion because the confidence interval is a way of estimating the mean of the population.  We would know if we took 100 estimates that the actual mean falls within the confidence intervals 95% of the time.  Within this context it is meaningless.*

@@

*Since there are so many variables I will be using a random forest to detect which variables are the most important and will be using them to build the model.  First I will drop out the variables that are missing a lot of data or otherwise meaningless.*

```{r rf, cache=TRUE, message=FALSE, warning=FALSE}
set.seed(12345)

rf_data <- kaggle %>%
  # Dropping variables with a lot of missing values
  select(-Alley, -MasVnrType, -MasVnrArea, -BsmtQual, -BsmtCond, -BsmtExposure, -BsmtFinType1, -BsmtFinType2, -Electrical, -FireplaceQu, -GarageType, -GarageYrBlt, -GarageFinish, -GarageQual, -GarageCond, -PoolQC, -Fence, -MiscFeature) %>%
  # Dropping the ID because it is meaningless
  select(-Id)

ctrl <- trainControl(method = "cv", number = 5)
rf_fit <- train(SalePrice ~ ., 
                data = rf_data, 
                method = "rf", 
                ntree = 1000, 
                trControl = ctrl, 
                tuneGrid = data.frame(mtry = 10), 
                importance = TRUE, 
                verbose = TRUE)
```

*Now we can examine the variables to in order of importance:*

```{r rf_importance}
imp_var <- importance(rf_fit$finalModel) 
variables <- rownames(imp_var)

@@

*The takeaways from the preceeding table is that size seems to consistently rise to the top as a variable to include in the model.  I don't want to have multiple variables that are measuring the same thing (i.e. Total SF of Living Area, Total SF of Basement, etc.).  I will use this table to select the variables that I think should be included in the linear regression model.*

```{r model, comment=NA}
fit <- lm(SalePrice ~ GrLivArea + GarageArea + LotArea + YearRemodAdd + Fireplaces + YearBuilt + OverallQual, kaggle)
summary(fit)
```

*These seven variables explain 80% of the variability in the sale price.  I will use this model to make the Kaggle submission.*

```{r predict, comment=NA}
test <- read.csv("Kaggle_House_Prices/test.csv")
y_hat <- predict(fit, test)
submission <- data.frame(Id = test$Id, SalePrice = y_hat)
summary(submission)
```

*It appears there is an NA in the submission.  In the interest of getting my final turned in on time I will be using the median to plug that hole.*

```{r}
fill <- median(submission$SalePrice, na.rm = TRUE)
submission <- submission %>%
  mutate(SalePrice = replace_na(SalePrice, fill))
write.csv(submission, "submission.csv", quote = FALSE, row.names =  FALSE)
```

*My Kaggle score is* **0.47768** *and my username is* **mikesilva***.  I am ranked 4520 on the leaderboard.  I will try to improve this score when I have time.*

*This analysis is available on  [GitHub](https://github.com/mikeasilva/CUNY-SPS/blob/master/DATA605/MSilva_Final.Rmd).*
